#lab_07 

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(tidyverse)
library(rvest)
library(janitor)
```


Q1. We are using NCAA Demographics Data showing race, gender for athletic directors and student athletes. ncaa_athletic_dirs.csv

Background on the data:
https://www.ncaa.org/sports/2018/12/13/ncaa-demographics-database.aspx?id=1729

Load the data into a dataframe called ads1. 
Create another dataframe where this data is filtered for 2021 only.
Using ggplot, create a basic column chart showing athletic directors by race and sex. 
Write a headline and caption.

**A1**

```{r}

#your answer here
ads1 <- read_csv("/Users/elajalil/Documents/GitHub/ela_jalil_jour472/lab_06/ncaa_athletic_dirs.csv") %>% 
  as.data.frame()
ads1_2021<-ads1 %>% 
  as.data.frame() %>% 
  filter(year==2021)

ads1_2021 %>% 
  ggplot(aes(x = race_sex, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Athletic Directors for the NCAA by Race and Sex", 
       subtitle = "NCAA data: 2021",
       caption = "Source: NCAA, Graphic by: Ela Jalil",
       y="Count",
       x="Sex and Race")
 

```

Q2. Now create a second chart using ads1 (ncaa_athletic_dirs) into a chart of columns with years on the x axis.
Write a headline and caption.

**A2**

```{r}
#your answer here  
ads1 %>% 
  ggplot(aes(x = year, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Athletic Directors for the NCAA from 2012-2021", 
       subtitle = "NCAA data, 2012-2021",
       caption = "Source: NCAA, Graphic by: Ela Jalil",
       y="Count",
       x="Year")


```

Q3. We will now visualize the data over the years for white men, black men and white women and black women athletic directors using this data: Sports_Race_ADs.csv

Remember to clean the names. Import the data with the appropriate tab and it beginning with the header row.

For example:
dataframe <- read_csv('my_spreadsheet.csv', skip =1)
Skip =1 allows the data to be imported beginning with the second row. This is done with some spreadsheets that have headlines and other junk in the first row.


Load the data, create a line chart showing athletic directors by race & sex. 
Designate one line for each and a separate color: ie male_white would be blue, female_other would be pink, etc. 

**A3.**

```{r}

#your answer here
sports_race <- read_csv('/Users/elajalil/Documents/GitHub/ela_jalil_jour472/lab_06/Sports_Race_ADs.csv', skip =1)
  ggplot(sports_race,aes(x=Year)) +
  geom_line(aes(y=`Male, White`), color="green")+
    geom_line(aes(y=`Male, Black`), color="yellow")+
    geom_line(aes(y=`Male, Other`), color="red")+
    geom_line(aes(y=`Female, White`), color="orange")+
    geom_line(aes(y= `Female, Black`), color="blue")+
    geom_line(aes(y=`Female, Other`), color="black")+
    labs(
      title="NCAA Directors by Sex and Race",
      x="Year",
      y= "Sex and Race"
    )

```
