#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for:

-   Total population
-   White only population
-   Black or African American only population
-   Hispanic only population
-   Native American only population
-   Asian only population
-   Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001)

```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

# your answer here
library(readr) #ask if this is necessary to include
lab_03 <- read_csv("lab_03.csv")
md_race <- lab_03
```

## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020.

A1:

```{r}

# Write your code in this codeblock.

# your answer here

md_race %>% 
  select(place,x2020_hispanic) %>% 
  slice_max(x2020_hispanic,n=5)
#The top five cities that have the largest Hispanic population in 2020 are Baltimore City (32,627), Wheaton CDP (22,787), Germantown CDP (21,381), Silver Spring CDP (21,087), Aspen Hill CDP (18,635).
```

Q2: Find the cities with the largest Black population and arrange the results in descending order.

A2:

```{r}
#write your code here

# your answer here
md_race %>% 
select(place,x2020_black) %>% 
  arrange(desc(x2020_black))
#The five cities with the largest Black population are Baltimore City (375,002), Waldorf CDP (47,903), Bowie City (32,784), Clinton CDP (30,679), Columbia CDP (28,991). 
```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor.

A3:

```{r}

# your answer here
top_hispanic <-md_race %>% 
 select(place,x2020_hispanic,x2010_hispanic,x2020_total) %>% 
slice_max(x2020_hispanic,n=15)
view(top_hispanic)

# Something that I noticed from this data is that seven of the 15 cities (47%) included in the table for the largest Hispanic population in 2020 were from Montgomery County: Wheaton CDP, Germantown CDP, Silver Spring CDP, Aspen Hill CDP, Gaitherburg City, Montgomery Village CDP and Rockville City. That leads me to wonder what attracts the Hispanic population to this county/cities - is it the location, resources, job opportunities or community? I would contact Marc Elrich to learn about how Montgomery County operates, and to see if there is anything in place on the executive level to attract a large Hispanic population into the area. 

#Baltimore County 1
#Montgomery County 7
#Prince George's County 4
#Frederick County 1
#Howard County 1
#Anne Arundel County 1
```
