#Lab_05  {lab_05}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)

```

### Load Life Expectancy Data

You'll need to load :

CDC_Life_Census_Tract_2010_2015.csv

It is included in your lab_05.zip file.

About the data: 
Life Expectancy at Birth for U.S. States and [Census Tracts, 2010-2015](https://www.cdc.gov/nchs/data-visualization/life-expectancy/index.html) 

It shows estimates of U.S. life expectancy at birth by state and census tract for the period 2010-2015. Estimates were produced for 67,148 census tracts, covering all states and the District of Columbia (D.C.). These estimates come from the National Center for Health Statistics (NCHS), the National Association for Public Health Statistics and Information Systems (NAPHSIS), and the Robert Wood Johnson Foundation (RWJF).


**Task** Create a codeblock below this task, then load the data sets and assign them to appropriate variable names. 

## Answer questions

**Q1.** In your lab_05 folder, there is a csv called CDC_Life_Census_Tract_2010_2015.csv.  It contains all of the life expectancy data for the United States.

You are going to examine this data and learn the range of death rates in Baltimore and then in Maryland in general.

Do the following:
-  Load tidyverse and janitor
-  Load the data CDC_Life_Census_Tract_2010_2015.csv into a dataframe called us_death. 
Include the code to show the number of records. Clean the names using Janitor
-  Filter this according to Maryland into second table called md_death
-  Filter this according to Baltimore City into a third table called balt_death

**A1.**

```{r}
library(tidyverse)
library(janitor)
us_death <- read_csv ("~/Documents/GitHub/ela_jalil_jour472/lab_05/CDC_Life_Census_Tract_2010_2015.csv") %>%
  as.data.frame()
#Number of records
nrow()
#Clean names
us_death_na <- na.omit(us_death)
#Create a second table
  md_death <- us_death_na %>% 
  filter(State=="Maryland")
#Create a third table
  balt_death <- md_death %>% 
    filter(County=="Baltimore city, MD")

```

**Q2.** 
For this question, load a modified dataset called balt_death_census.csv.

Call it balt_death2

This has the same information but with the Baltimore neighborhoods. 

Produce summary statistics showing the data distribution by quartile, median, mean, miniumum and maximum values on the life expectancy rate.

Show the neighborhoods with a life expectancy rate below the 1st quartile, arranged in ascending order. 


**A2.** 
```{r}
balt_death2 <- read_csv("~/Documents/GitHub/ela_jalil_jour472/lab_05/balt_death_census.csv") 
  #summary statistics
```
```{r}
view(balt_death2)
summary(balt_death2$life_expectancy)
balt_death2 %>%
  select(neighborhood, life_expectancy) %>% 
  filter(life_expectancy <69.58) %>% 
arrange(life_expectancy)
```

**Q3.** Use md_death, the Maryland life expectancy rates.
Produce summary statistics of the state of Maryland life expectancy rate. 
Produce a table with the 20 communities with the lowest life expectancy. 
Produce a table with the 20 communities with the highest life expectancy. 

Then use us_death, and determine the average life expectancy nationwide.

Then write 150-200 words on this document below your code that compares your findings of Baltimore, the state and nation. 

**A3.**

```{r}
#Average life expectancy for the US
summary(md_death$'Life Expectancy') 
md_death %>% 
slice_min(`Life Expectancy`, n=20) 
md_death %>% 
  slice_max(`Life Expectancy`,n=20)
#Average life expectancy for the US
us_death %>% 
  summary(us_death$`Life Expectancy`)

#The average life expectancy in Baltimore is 73.04 years, which is far behind the Maryland average life expectancy of 78.7 years, and the US average life expectancy of 78.5 years. While the difference between the Maryland and US average is minimal, there is a significant difference between Maryland and Baltimore’s life expectancy. I noticed that when I found the communities with the lowest life expectancy in Maryland, they were all from Baltimore city, with the lowest life expectancy rate being 62.6 years. I noticed that the highest life expectancy rates were from Montgomery County, with the highest life expectancy rate being 96.1 years. I think that it is interesting that there is such a large discrepancy between the counties in Maryland in regards to average life expectancy, and would be interested to look further into each county in Maryland and see the impact of resources and demographics on the average life expectancy. 

```

**Q4.**

Use balt_death. Use mutate and the case_when command.

Categorize all Baltimore City Census tracts as being above or below the city's average life expectancy rate. 

Then produce a table that summarizes the number of Census tracts above or below the average life expectancy rate. Write two sentences summarizing your findings.

Pro Tip: See this chapter section for a refresher on case_when: https://wellsdata.github.io/data_journalism_class/_book/r-mutating.html#a-more-powerful-use


```{r}
death <- balt_death %>%  
mutate(
  category = case_when(
    `Life Expectancy` >= 73.04 ~ "above",
    `Life Expectancy` <= 73.04~ "below"
  )) %>% 
  select(`Census Tract Number`, `Life Expectancy`, category) 
value_counts <- death %>%
  count(category)
  print(value_counts)

# There are 87 census tracts that have life expectancy's above the average life expectancy and 93 census tracts below the average life expectancy. The life expectancy in Baltimore is far lower than the United States average life expectancy, and there are evidently many census tracts that are even lower than that. 

```


