---
editor: 
  markdown: 
    wrap: 72
---

#lab_07 Sean Mussenden

## About this lab

To complete this lab, you need to: \* write code in empty codeblocks
provided to answer questions included (look for **Q**). \* write out the
answer in the form of a complete sentence in the space given (look for
**A**).

When you are finished, commit changes and push to your personal GitHub
repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task**: Make a code block and load janitor and the tidyverse

```{r}
library(tidyverse)
library(janitor)
```

## Load data

For today's lab, we are using the DEA's ARCOS database, which tracks
shipments of opioids and other controlled substances from manufacturers
to pharmacies. As we will read this week, the Washington Post used this
data to show which communities were flooded by opioids during the death
and overdose crisis of the last two decades.

There are four tables in the data folder.

-   arcos.csv has the total number of opioid pills shipped to pharmacies
    located in each U.S. county each year (from 2006-2014). Each row
    represents one county in one year during that period. In some
    states, counties and cities are represented. Each county has a
    unique FIPS code, which stands for Federal Information Processing
    System (FIPS), which identify geographic areas. You will need this
    for both questions.
-   county_population_2014.csv has the total population of each U.S.
    county (or city in some states) in 2014 from the American Community
    Survey. It has a unique numeric identifier for the county called
    GEOID. You will need to use this for the first question.
-   state_population_2014.csv has the total population of each U.S.
    state in 2014 from the American Community Survey. You will need to
    use this for the second question.
-   state_crosswalk.csv is a simple dataframe with two columns listing
    the state and its abbreviation. You may need to use this for the
    second question.

**Task**: Create a codeblock and load the four dataframes using
appropriate names below.

```{r}
arcos <- read_csv("data/arcos.csv") %>% 
  clean_names()
county_pop2014 <- read_csv("data/county_population_2014.csv") %>%
   clean_names()
state_pop2014 <- read_csv("data/state_population_2014.csv") %>% 
  clean_names()
state_crosswalk <- read_csv("data/state_crosswalk.csv") %>% 
  clean_names()
```

## Questions

**Task** Create codeblocks in the space below and answer the questions.

**Q1** Which county or city received the most total opioid pills per
person? Add up all the pills between 2006 and 2014 when doing this
calculation. How many total pills, and what was the per person rate?

help(group_by)

```{r}
#This adds up all of the pills between 2006 and 2014 by county

arcos_county <- arcos %>%
  group_by(countyfips,buyer_county,buyer_state) %>% 
  summarise(
    total_pills = sum(total_pills)
    ) %>% 
  inner_join(county_pop2014, by=c("countyfips"="geoid")) %>%
  mutate(pills_per_person = total_pills/total_population_2014) %>% 
  arrange(desc(pills_per_person))

arcos_county

#Norton City, Virginia receieved the most total pills per person. Its total amount of pills is 10555170 with a total population in 2014 of 4015. This makes its pills per person rate 2628.9340. 
```

**Q2_New** Which U.S. state saw the highest percentage increase in opiod
pills from 2006 to 2014?

--Produce two tables, one for 2006 and one for 2014. In the 2006 table,
rename the total to total_2006. In the 2014 table, rename the total to
total_2014. --Join the tables into a new table, arcos_combo --Your table
must have the full name of the states. Join with the state crosswalk
table --Calculate the percentage change between 2006 and 2014 --Create
one table with the top 5 states by percentage increase --Create a second
table with the 5 states with the lowest percentage increase.

--Write a sentence that could be dropped in as the lede of a news story
from the finding in question two.

```{r}
#table one
total_2006 <-arcos %>% 
  filter(year==2006) %>% 
  clean_names()
#table two
total_2014 <- arcos %>% 
  filter(year==2014) %>% 
  clean_names()
#combining 2006 and 2014
arcos_combo <- total_2006 %>%
  inner_join(total_2014, by="countyfips") %>% 
  clean_names()
arcos_combo_sidewalk<- arcos_combo %>% inner_join(state_crosswalk, by=c("buyer_state_x"="state_abb"))
new_arcos <-arcos_combo_sidewalk %>% 
  select(state_name, total_pills_x, total_pills_y) %>% 
  mutate(pct_change = ((total_pills_y-total_pills_x)/(total_pills_x)*100)) 
top_5 <- new_arcos %>%
  slice_max(pct_change, n=5)
bottom_5 <- new_arcos %>% 
  slice_min(pct_change, n=5)
# The top 5 states with a percentage increase between 2006 and 2014 were Missouri (36118.9055%), Alaska (4550.0000%), Texas (2500.0000%), Virginia (913.2187%), and South Dakota (434.5588%).

#Lede: Missouri had the largest increase in total opioid pills between 2006 and 2014, an analysis of county opioid data shows. 
```

**Q3** Which U.S. state had the highest number of opioid pills per
person for pills shipped in 2014? How many total pills, and what was the
per person rate? How did that compare to Washington, D.C.?

```{r}

```
